Paper_Cartama_lenticel_damage

### 2023-04-01 R/Paper_plots_lenticel_damage.Rmd
before last revision

```{r Packages, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(gridExtra)
library(readxl)
library(ggpubr)
library(tibble)
library(ggplot2)
library(extrafont)
loadfonts(device = "win")

@@

df <- Lenticel_file %>% drop_na(Severidad_0 , Incidencia_0)
colnames(df)[which(names(df) == "Cosecha")] <- "Harvest"
colnames(df)[which(names(df) == "UP")] <- "Farm"
colnames(df)[which(names(df) == "Lote")] <- "Plot"
colnames(df)[which(names(df) == "Severidad_0")] <- "Sev_0dph"
colnames(df)[which(names(df) == "Severidad_1")] <- "Sev_21dph"
colnames(df)[which(names(df) == "Incidencia_0")] <- "Inc_0dph"
colnames(df)[which(names(df) == "Incidencia_1")] <- "Inc_21dph"
glimpse(df)
summary(df$Sev_0dph)

@@

for (i in 2:13){
  sev<-c(sev , df_mean[i,])
  inc <- c(inc , df_mean1[i,])
  }
mat <- data.frame(rep(c("2019_m" , "2020_m" , "2020_t" , "2021_t") , 13) , c(rep("Escondida" , 24) , rep("Sinai" , 28)) , sev , inc)

@@

p <-  ggplot(mat, aes(x = factor(Harvest,levels = c("2019_m","2020_t","2020_m","2021_t")), y = Severity)) + 
      geom_bar(stat="summary", color="black", position=position_dodge(), fill=rep(c("gray90","gray70"),4),fun = mean) +
      stat_summary(fun.data=mean_sdl, fun.args = list(mult=1), geom="errorbar", color="black", width=0.2) 
p <- p + geom_jitter()+ facet_wrap(~Farm) + theme_classic() 
p <- p +  xlab("") + ylab("Necrotic Area (%)")
p <- p + theme(axis.title =element_text(size=14,family = 'serif'))
p <- p +  theme(axis.text.x = element_text(face="bold",size=12, angle=0,family = 'serif'),
              axis.text.y = element_text(face="bold",size=12, angle=0,family = 'serif'))
p <- p + theme(strip.text.x = element_text(size = 14,family = 'serif'))

@@

p1 <-  ggplot(mat, aes(x = factor(Harvest,levels = c("2019_m","2020_t","2020_m","2021_t")), y = Incidence)) + 
      geom_bar(stat="summary", color="black", position=position_dodge(), fill=rep(c("gray90","gray70"),4),fun = mean) +
      stat_summary(fun.data=mean_sdl, fun.args = list(mult=1), geom="errorbar", color="black", width=0.2) 
p1 <- p1 + geom_jitter()+ facet_wrap(~Farm) + theme_classic() 
p1 <- p1 +  xlab("") + ylab("Necrotic spots")
p1 <- p1 + theme(axis.title =element_text(size = 14 , family = 'serif'))
p1 <- p1 +  theme(axis.text.x = element_text(face="bold", size = 12, angle = 0 , family = 'serif'),
              axis.text.y = element_text(face = "bold" , size = 12 , angle = 0 , family = 'serif'))
p1 <- p1 + theme(strip.text.x = element_text(size = 14 , family = 'serif'))

@@

 plot<-grid.arrange(p + labs(tag = 'A'),p1 + labs(tag = 'B'), ncol=1,nrow=2)
ggsave("C:/Users/Dru/Desktop/Cartama_paper/Figure1.png",plot=plot,dpi = 300,width = 7,height = 7,units = c("in"))

@@

sev <- c(df1$Sev_0dph , df1$Sev_21dph)
inc <- c(df1$Inc_0dph , df1$Inc_21dph)
farm <- c(df1$Farm , df1$Farm)
time <-factor(c(rep("0dpc",length(sev)/2) , rep("21dpc",length(sev)/2)))
harvest <- c(df1$Harvest , df1$Harvest)

@@

```{r severity in time}
p <-  ggplot(mat1, aes(x = factor(Time), y = Severity)) + 
      geom_bar(stat="summary", color="black", positio=position_dodge(), fill= rep(c("gray95","gray85") ,2), fun = mean) +
      stat_summary(fun.data=mean_sdl, fun.args = list(mult=1), geom="errorbar", color="black", width=0.2) 
p <- p + geom_jitter(aes(colour = factor(Harvest , levels = c("2019_m" , "2020_t" , "2020_m" , "2021_t"))) , size = 2)
p <- p + facet_wrap(~ Farm) + theme_classic() 
p <- p +  xlab("") + ylab("Necrotic Area (%)")
p <- p + theme(axis.title =element_text(size=14,family = 'serif'))
p <- p +  theme(axis.text.x = element_text(face="bold",size=12, angle=0,family = 'serif'),
              axis.text.y = element_text(face="bold",size=12, angle=0,family = 'serif'))
p <- p + theme(strip.text.x = element_text(size = 14,family = 'serif'))
p <- p + scale_color_manual(name = "Harvest", values =  rep(c("gray70","gray40") ,2))
p <- p + theme(legend.text = element_text(size=12, angle=0,family = 'serif'),
                legend.title = element_text(face="bold",size=12, angle=0,family = 'serif'))

@@

```{r incidence in time}
p1 <-  ggplot(mat1, aes(x = factor(Time), y = Incidence)) + 
      geom_bar(stat="summary", color="black", positio=position_dodge(), fill= rep(c("gray95","gray85") ,2), fun = mean) +
      stat_summary(fun.data=mean_sdl, fun.args = list(mult=1), geom="errorbar", color="black", width=0.2) 
p1 <- p1 + geom_jitter(aes(colour = factor(Harvest , levels = c("2019_m" , "2020_t" , "2020_m" , "2021_t"))) , size = 2)
p1 <- p1 + facet_wrap(~ Farm) + theme_classic() 
p1 <- p1 +  xlab("") + ylab("Necrotic spots")
p1 <- p1 + theme(axis.title =element_text(size=14,family = 'serif'))
p1 <- p1 +  theme(axis.text.x = element_text(face="bold",size=12, angle=0,family = 'serif'),
              axis.text.y = element_text(face="bold",size=12, angle=0,family = 'serif'))
p1 <- p1 + theme(strip.text.x = element_text(size = 14,family = 'serif'))
p1 <- p1 + scale_color_manual(name = "Harvest", values =  rep(c("gray70","gray40") ,2))
p1 <- p1 + theme(legend.text = element_text(size=12, angle=0,family = 'serif'),
                legend.title = element_text(face="bold",size=12, angle=0,family = 'serif'))

@@

 plot<-grid.arrange(p + labs(tag = 'A'),p1 + labs(tag = 'B'), ncol=1,nrow=2)
ggsave("C:/Users/Dru/Desktop/Cartama_paper/Figure2.png",plot=plot,dpi = 300,width = 7,height = 7,units = c("in"))

>>>>

```{r Packages, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(gridExtra)
library(readxl)
library(ggpubr)
library(tibble)
library(ggplot2)
library(extrafont)
loadfonts(device = "win")
library(ggsignif)
library(ggpubr)

@@

df <- Lenticel_file %>% drop_na(Severidad_0 , Incidencia_0)
colnames(df)[which(names(df) == "Cosecha")] <- "Harvest"
colnames(df)[which(names(df) == "UP")] <- "Farm"
colnames(df)[which(names(df) == "Lote")] <- "Plot"
colnames(df)[which(names(df) == "Severidad_0")] <- "Sev_0dph"
colnames(df)[which(names(df) == "Severidad_1")] <- "Sev_21dph"
colnames(df)[which(names(df) == "Incidencia_0")] <- "Inc_0dph"
colnames(df)[which(names(df) == "Incidencia_1")] <- "Inc_21dph"
df$Inc_increase <- df$Inc_21dph-df$Inc_0dph
df$Sev_increase <- df$Sev_21dph-df$Sev_0dph

glimpse(df)
summary(df$Sev_0dph)

@@

for (i in 2:13){
  sev<-c(sev , df_mean[i,])
  inc <- c(inc , df_mean1[i,])
  }
mat <- data.frame(rep(c("2019_m" , "2020_m" , "2020_t" , "2021_t") , 13) , c(rep("Escondida" , 20) , rep("Sinai" , 32)) , sev , inc)

@@

p <-  ggplot(mat, aes(x = factor(Harvest,levels = c("2019_m","2020_t","2020_m","2021_t")), y = Severity)) + 
      geom_bar(stat="summary", color="black", position=position_dodge(), fill=rep(c("gray90","gray70"),4),fun = mean) +
      stat_summary(fun.data=mean_sdl, fun.args = list(mult=1), geom="errorbar", color="black", width=0.2) 
p <- p + geom_jitter()+ facet_wrap(~Farm) + theme_classic() 
p <- p +  xlab("") + ylab("Necrotic Area (%)")
p <- p + theme(axis.title =element_text(size=14,family = 'serif'))
p <- p +  theme(axis.text.x = element_text(face="bold",size=12, angle=0,family = 'serif'),
              axis.text.y = element_text(face="bold",size=12, angle=0,family = 'serif'))
p <- p + theme(strip.text.x = element_text(size = 14,family = 'serif'))
p

@@

p1 <-  ggplot(mat, aes(x = factor(Harvest,levels = c("2019_m","2020_t","2020_m","2021_t")), y = Incidence)) + 
      geom_bar(stat="summary", color="black", position=position_dodge(), fill=rep(c("gray90","gray70"),4),fun = mean) +
      stat_summary(fun.data=mean_sdl, fun.args = list(mult=1), geom="errorbar", color="black", width=0.2) 
p1 <- p1 + geom_jitter()+ facet_wrap(~Farm) + theme_classic() 
p1 <- p1 +  xlab("") + ylab("Number of necrotic spots")
p1 <- p1 + theme(axis.title =element_text(size = 14 , family = 'serif'))
p1 <- p1 +  theme(axis.text.x = element_text(face="bold", size = 12, angle = 0 , family = 'serif'),
              axis.text.y = element_text(face = "bold" , size = 12 , angle = 0 , family = 'serif'))
p1 <- p1 + theme(strip.text.x = element_text(size = 14 , family = 'serif'))
p1

@@

 plot<-grid.arrange(p + labs(tag = 'A'),p1 + labs(tag = 'B'), ncol=1,nrow=2)
ggsave("C:/Users/Dru/Desktop/Cartama_paper/Figure1_new.jpg",plot=plot,dpi = 300,width = 7,height = 7,units = c("in"))

@@

sev <- c(df1$Sev_0dph , df1$Sev_21dph)
inc <- c(df1$Inc_0dph , df1$Inc_21dph)

###stack plot
#sev <- c(df1$Sev_0dph , df1$Sev_increase)
#inc <- c(df1$Inc_0dph , df1$Inc_increase)
farm <- c(df1$Farm , df1$Farm)
time <-factor(c(rep("0dpc",length(sev)/2) , rep("21dpc",length(sev)/2)))
harvest <- c(df1$Harvest , df1$Harvest)

@@

```{r severity in time}
p <-  ggplot(mat1, aes(x = factor(Time), y = Severity)) + 
      geom_bar(stat="summary", color="black", positio=position_dodge(), fill= rep(c("gray99","gray95") ,2), fun = mean) +
      stat_summary(fun.data=mean_sdl, fun.args = list(mult=1), geom="errorbar", color="black", width=0.2) 
p <- p + geom_jitter(aes(colour = factor(Harvest , levels = c("2019_m" , "2020_t" , "2020_m" , "2021_t"))) , size = 2)
p <- p + facet_wrap(~ Farm) + theme_classic() 
p <- p +  xlab("") + ylab("Necrotic Area (%)")
p <- p + theme(axis.title =element_text(size=14,family = 'serif'))
p <- p +  theme(axis.text.x = element_text(face="bold",size=12, angle=0,family = 'serif'),
              axis.text.y = element_text(face="bold",size=12, angle=0,family = 'serif'))
p <- p + theme(strip.text.x = element_text(size = 14,family = 'serif'))
p <- p + scale_color_manual(name = "Harvest",  values =  c("gray88","gray70","gray47","gray0"))
p <- p + theme(legend.text = element_text(size=12, angle=0,family = 'serif'),
                legend.title = element_text(face="bold",size=12, angle=0,family = 'serif'))
p

@@

```{r incidence in time}
p1 <-  ggplot(mat1, aes(x = factor(Time), y = Incidence)) + 
      geom_bar(stat="summary", color="black", positio=position_dodge(), fill= rep(c("gray99","gray95") ,2), fun = mean) +
      stat_summary(fun.data=mean_sdl, fun.args = list(mult=1), geom="errorbar", color="black", width=0.2) 
p1 <- p1 + geom_jitter(aes(colour = factor(Harvest , levels = c("2019_m" , "2020_t" , "2020_m" , "2021_t"))) , size = 2)
p1 <- p1 + facet_wrap(~ Farm) + theme_classic() 
p1 <- p1 +  xlab("") + ylab("Number of necrotic spots")
p1 <- p1 + theme(axis.title =element_text(size=14,family = 'serif'))
p1 <- p1 +  theme(axis.text.x = element_text(face="bold",size=12, angle=0,family = 'serif'),
              axis.text.y = element_text(face="bold",size=12, angle=0,family = 'serif'))
p1 <- p1 + theme(strip.text.x = element_text(size = 14,family = 'serif'))
p1 <- p1 + scale_color_manual(name = "Harvest", values =  c("gray88","gray70","gray47","gray0"))
p1 <- p1 + theme(legend.text = element_text(size=12, angle=0,family = 'serif'),
                legend.title = element_text(face="bold",size=12, angle=0,family = 'serif'))

@@

 plot<-grid.arrange(p + labs(tag = 'A'),p1 + labs(tag = 'B'), ncol=1,nrow=2)
ggsave("C:/Users/Dru/Desktop/Cartama_paper/Figure2_new.jpg",plot=plot,dpi = 300,width = 7,height = 7,units = c("in"))

```
```{r df_for_barplot plots}

sev <- c(df1$Sev_0dph , df1$Sev_21dph)
inc <- c(df1$Inc_0dph , df1$Inc_21dph)

farm <- c(df1$Farm , df1$Farm)
time <-factor(c(rep("0dph",length(sev)/2) , rep("21dph",length(sev)/2)))
harvest <- c(df1$Harvest , df1$Harvest)

mat_mean<-tapply(mat$Severity , list(mat$Time,mat$Harvest,mat$Farm),mean)
mat_mean1<-tapply(mat$Incidence , list(mat$Time,mat$Harvest,mat$Farm),mean)
mat_se<-tapply(mat$Severity , list(mat$Time,mat$Harvest,mat$Farm),sd)/sqrt(table(mat$Time,mat$Harvest,mat$Farm))
mat_se1<-tapply(mat$Incidence , list(mat$Time,mat$Harvest,mat$Farm),sd)/sqrt(table(mat$Time,mat$Harvest,mat$Farm))
n <- table(mat$Time,mat$Harvest,mat$Farm)

mat1<-data.frame(Farm = c(rep("Escondida" , 8) , rep("Sinai" , 8)), Time = rep(c(rep("0dph",4) , rep("21dph",4)) , 2) ,
                 Harvest = rep(c("2019_m" , "2020_m" , "2020_t" , "2021_t") , 4), 
                 Severity = c(mat_mean[1,,1] , mat_mean[2,,1] , mat_mean[1,,2] , mat_mean[2,,2]),
                 Severity_se = c(mat_se[1,,1] , mat_se[2,,1] , mat_se[1,,2] , mat_se[2,,2]),
                 Incidence = c(mat_mean1[1,,1] , mat_mean1[2,,1] , mat_mean1[1,,2] , mat_mean1[2,,2]),
                 Incidence_se = c(mat_se1[1,,1] , mat_se1[2,,1] , mat_se1[1,,2] , mat_se1[2,,2]))

```{r severity in time}

data_hline <- data.frame(Farm = c(rep("Escondida",4),rep( "Sinai",4)),
                         x = c(0.7,1.7,2.7,3.7,0.7,1.7,2.7,3.7), y = c(1.2, 2.4,0.9,4.7,2.8,4.6,5.4,3.6),xend = c(1.3,2.3,3.3,4.3,1.3,2.3,3.3,4.3), yend = c(1.2, 2.4,0.9,4.7,2.8,4.6,5.4,3.6), label = rep("*",8))

p <-  ggplot(mat1, aes(x = factor(Harvest, levels = c("2019_m" , "2020_t" , "2020_m" , "2021_t")), y = Severity, fill = factor(Time))) + geom_bar(stat="identity", positio=position_dodge())     
p <- p + scale_fill_manual(values=c("gray90", "gray80"))
p <- p + geom_errorbar(aes(ymin=Severity-Severity_se, ymax=Severity+Severity_se), width=.2,position=position_dodge(.9)) 
p <- p + facet_grid(~ Farm) + theme_classic() 
p <- p +  xlab("") + ylab("Necrotic Area (%)") + labs(fill='Days postharvest') 
p <- p + theme(axis.title =element_text(size=14,family = 'serif'))
p <- p +  theme(axis.text.x = element_text(face="bold",size=12, angle=0,family = 'serif'),
              axis.text.y = element_text(face="bold",size=12, angle=0,family = 'serif'))
p <- p + theme(strip.text.x = element_text(size = 14,family = 'serif'))
p <- p + theme(legend.text = element_text(size=12, angle=0,family = 'serif'),
                legend.title = element_text(face="bold",size=12, angle=0,family = 'serif'))

p <- p + geom_segment(data = data_hline,aes(x = x,y = y,xend = xend, yend = yend),inherit.aes = FALSE) 
p <- p +geom_text(data = data_hline,aes(label = label,x = x+0.3,y = y+0.2),inherit.aes = FALSE)

```{r incidence in time}

p1 <-  ggplot(mat1, aes(x = factor(Harvest, levels = c("2019_m" , "2020_t" , "2020_m" , "2021_t")), y = Incidence, fill = factor(Time))) + geom_bar(stat="identity", positio=position_dodge())     
p1 <- p1 + scale_fill_manual(values=c("gray90", "gray80"))
p1 <- p1 + geom_errorbar(aes(ymin=Incidence-Incidence_se, ymax=Incidence+Incidence_se), width=.2,position=position_dodge(.9)) 
p1 <- p1 + facet_wrap(~ Farm) + theme_classic() 
p1 <- p1 +  xlab("Harvest") + ylab("Number of necrotic spots") + labs(fill='Days postharvest') 
p1 <- p1 + theme(axis.title =element_text(size=14,family = 'serif'))
p1 <- p1 +  theme(axis.text.x = element_text(face="bold",size=12, angle=0,family = 'serif'),
              axis.text.y = element_text(face="bold",size=12, angle=0,family = 'serif'))
p1 <- p1 + theme(strip.text.x = element_text(size = 14,family = 'serif'))
p1 <- p1 + scale_color_manual(name = "Harvest", values =  c("gray88","gray70","gray47","gray0"))
p1 <- p1 + theme(legend.text = element_text(size=12, angle=0,family = 'serif'),
                legend.title = element_text(face="bold",size=12, angle=0,family = 'serif'))

data_hline <- data.frame(Farm = c(rep("Escondida",4),rep( "Sinai",4)),
                         x = c(0.7,1.7,2.7,3.7,0.7,1.7,2.7,3.7), y = c(88, 165,70,203,238,217,256,187),xend = c(1.3,2.3,3.3,4.3,1.3,2.3,3.3,4.3), yend = c(88, 165,70,203,238,217,256,187), label = rep("*",8))

p1 <- p1 + geom_segment(data = data_hline,aes(x = x,y = y,xend = xend, yend = yend),inherit.aes = FALSE) 
p1 <- p1 + geom_text(data = data_hline,aes(label = label,x = x+0.3,y = y+5),inherit.aes = FALSE)
p1

```
*Fig 2.* Lenticel damage for avocado cv. Hass fruits collected from the La Escondida and El Sinai farms between 2019 and 2021 at harvest time (0 days post-harvest, 0 dph) and after 21 days of storage at 6C (21 days post-harvest, 21 dph). Shown are the means and standard errors for the severity (A) and incidence (B) of the damage for the two evaluatin points (0 dph and 21 dph). The * denotes statistical differences at a 95% confidence level according to the  generalized linear mixed models, and the letter after the year denotes main (m) and traviesa (t) harvest.

 plot<-grid.arrange(p + labs(tag = 'A'),p1 + labs(tag = 'B'), ncol=1,nrow=2)
ggsave("C:/Users/Dru/Desktop/Cartama_paper/Figure1_new.jpg",plot=plot,dpi = 300,width = 7,height = 7,units = c("in"))
